### 2021-03-01 paper/paper.tex
Cleaned code; added docstrings.

\documentclass{iopart}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric}
\usepackage{bbm}
\expandafter\let\csname equation*\endcsname\relax
\expandafter\let\csname endequation*\endcsname\relax
\usepackage{amsmath}
\newcommand{\pd}[2]{\frac{\partial#1}{\partial#2}}
\newcommand{\undertilde}[1]{\underset{\tilde{}}{#1}}
\newcommand{\hmatrix}{\ensuremath{%
  \begin{pmatrix}
    \mathbbm{O}&\mathbbm{I}\\-\mathbbm{I}&\mathbbm{O}
  \end{pmatrix}
}}
\newcommand{\numap}[1]{\ensuremath{%
    \left\lfloor#1\right\rceil
}}

\begin{document}
\title{Numerical Evolution from the Hamiltonian}
\author{José M L Amoreira$^1$ and Luís J M Amoreira$^2$}
\address{$^1$ Departamento de Física, Instituto Superior Técnico, Lisboa,
Portugal}
\address{$^2$ Departamento de Física, Universidade da Beira Interior, Covilhã,
Portugal}
\ead{amoreira@ubi.pt}
\begin{abstract}
  We propose a numerical method for approximate calculations of the time
  evolution of point particle systems given only the system's hamiltonian
  function and initial conditions. The method both generates and solves the
  equations of motion numerically. For demonstration purposes, a working
  im\-ple\-men\-ta\-tion (written in \texttt{python}) is described and applied
  to standard problems. The method may have some pedagogical merits but the
  numerical effort of generating the equations of motion makes it unsuitable for
  actual numerical solution of ``real'' problems with any but just a few degrees
  of freedom.
\end{abstract}
\noindent{\it Keywords\/}: Classical particle dynamics,
Hamilton's equations,
Numerical methods

\submitto{\EJP}
%\maketitle
\section{Introduction}
%-----
% 1 As equações de Hamilton de um sistema de massas pontuais formam um sistema
% de N ODEs de 1ª ordem, com a forma 
% \dot x_i= G_i(t, \tilde x)
%-----
In the hamiltonian formulation of classical mechanics~\cite{gol:1980, fw:2003},
the time evolution of physical systems is governed by Hamilton's equations
which, for point particle systems, take the form
\begin{align}\label{eq:heqs}
  \dot q_i&=\pd{H}{p_i}&
  \dot p_i&=-\pd{H}{q_i},&
  i&=1, 2, \ldots, N,
\end{align}
where $N$ is the number of degrees of freedom of the system (three times the
number of particles for three dimensional, unconstrained systems), $q_i$, $p_i$
are the canonical coordinates and conjugate momenta respectively, $H=H(t, q_i,
p_i)$ is the hamiltonian function of the system (for systems with time
independent potentials and constraints, the hamiltonian is simply their
mechanical energy function~\cite{fw:2003}) and dotted symbols denote their
total derivative with respect to time. 
Renaming the canonical coordinates and momenta as
\begin{equation}
  x_j = 
  \begin{cases}
    q_i&\quad\text{if }j\leq N\\
    p_{i-N}&\quad\text{if }j> N
  \end{cases}
  \qquad j=1, \ldots, 2N,
\end{equation}
Hamilton's equations take a more unified form
\begin{equation}\label{eq:std_ode}
  \dot x_j=G_j(t, x_1, x_2, \ldots, x_{2N}),\quad j=1, \ldots, 2N
\end{equation}
where
\begin{equation}
  G_j(t, x)=
  \begin{cases}
    \frac{\partial H}{\partial x_{j+N/2}},&\quad\text{if }j\leq N\\
    \frac{\partial H}{\partial x_{j-N/2}},&\quad\text{if }j> N.
  \end{cases}
\end{equation}
This expression can be further simplified as a matrix product:
\begin{equation}\label{eq:rhsmf}
  G_i=M_{i j}\frac{\partial H}{\partial x_j},\qquad
  \text{with }
  M=
  \begin{pmatrix}
    \mathbbm{O}&\mathbbm{I}\\
    -\mathbbm{I}&\mathbbm{O}
  \end{pmatrix},
\end{equation}
where $\mathbbm{O}$ and $\mathbbm{I}$ represent, respectively, the $N\times N$
zero and identity matrices.  Equation~\eqref{eq:std_ode} shows the Hamilton's
equations for a system of point particles with $N$ degrees of freedom as a
system of $2N$ first order ordinary differential equations (ODEs) on $2N$
variables $x_1, x_2, \ldots, x_{2N}$.

In all but a small handfull of very well known simple problems, this system of
ODEs must be solved numerically. All popular library routines for solving
ODEs\footnote{Like ODEPACK \cite{odepack} for fortran, \texttt{odeint}
  \cite{odeint} in c++ boost libraries or 
  \texttt{scipy.integrate.solve\_ivp}~\cite{scipy} for python.}
require the user to supply subprograms to compute the rhs functions $G_i$.
The ODE solver invokes these subprograms to compute estimates for the
values of the unknowns $x_i$ for arbitrary time $t$, given their values at a
given instant $t_0$.

In this work, we propose an alternative numerical method, where the user only
supplies the hamiltonian function of the system, leaving its partial derivatives
to be computed numerically as well.  This procedure is computationally more
expensive, since many evaluations of the hamiltonian function are needed in
order to estimate its partial derivatives at any moment in time, and so it
really is not suitable for the analysis of complex systems with more than just a
few coupled degreess of freedom. However, for simple systems it is very
practical and, anyway, it is an interesting concept in its own, which we haven't
seen exposed elsewhere.

%-------------------------------------------------------------------------------
\section{Numerical procedure}
Standard numerical routines for solving systems of ODEs expect the problem to be
written in the form
\begin{equation}\label{eq:numheqs}
  \frac{d\psi_i}{dt}=G_i(t, \psi_1, \psi_2, \ldots),\quad i=1, \ldots, N,
\end{equation}
where $N$ is the number of unknown functions $\psi_i(t)$ (and the number of
equations in well posed problems) and $G_i(t, \psi_1, \psi_2, \ldots)\equiv
G_i(t,\psi)$ are the rhs of the differential equations, wich must be coded by
the user as callable subprograms. For instance, the python standard initial
value problem solver routine has interface
\texttt{solve\_ivp(rhs, time\_interval, psi\_0)}; here, \texttt{rhs} is the name
of a phython function to compute the values of the rhs functions $G_i$.
During its execution, the ODE solver routine repeatedly invokes these
subprograms (as many times as required by the specific integration algorithm
and/or desired accuracy) in order to compute estimates for the values the
unknown functions $\psi_i$ at given values of $t$.

For Hamilton's equations, the $\psi_i$ are the coordinates and conjugate momenta
of the system, while the $G_i$ are the partial derivatives of its hamiltonian
function (reordered and eventually symmetrized, according to
eq.~\ref{eq:rhsmf}). The usual procedure is to derive analytical expressions
for the $G_i$ as functions of $t$, the coordinates and the momenta, and then to
code those expressions as subprograms to be called by the ODE solver routine.

In the method now proposed, instead, the user only has to supply code to
compute the value of the system's hamiltonian function; its partial derivatives
(the $G_i$ in eq.~\eqref{eq:numheqs}) are estimated numerically and reordered by
another subprogram which the ODE solver calls. Note that the differentiation and
reordering of the partial derivatives can be computed by a general purpose
routine. Indeed, the gradient can even be computed also by a standard numerical
library function, if available for the chosen computing system or language.
Figure~\ref{fig:a} displays flowcharts for both methods.
\begin{figure}[htb]
  \centering
    \begin{tikzpicture}[node distance=1.0cm, font=\sffamily,scale=0.7]
      \footnotesize
      \tikzstyle{io} = [trapezium,
                        trapezium left angle=70,
                        trapezium right angle=110,
                        minimum width=1.0cm,
                        minimum height=.6cm,
                        text centered,
                        draw=black,
                        fill=blue!20]
      \tikzstyle{libcode} = [rectangle,
                             rounded corners,
                             minimum width=1.5cm,
                             text width=1.5cm,
                             minimum height=0.6cm,
                             text centered,
                             draw=black,
                             fill=orange!20]
      \tikzstyle{usercode} = [rectangle,
                             %rounded corners,
                             minimum width=1.5cm,
                             minimum height=0.6cm,
                             text centered,
                             draw=black,
                             text width=1.5cm,
                             fill=green!20]
      \tikzstyle{arrow} = [thick,->,>=stealth]

      \begin{scope}[xshift=-3.5cm]
        \node at (0,1) {Standard method};
        \node (input) [io] {Initial state};
        \node (odesolver) [libcode, below of=input] {ODE Solver};
        \node (output) [io, below of=odesolver] {Final state};
        \node (rhs) [usercode, right of=odesolver, node distance=2.3cm]
          {$M\,\text{grad}\,H$};
        \draw [arrow] (input) -- (odesolver);
        \draw [arrow] (odesolver) -- (output);
        \draw [arrow] (odesolver) -- (rhs);
        \draw [arrow]  (rhs) -- (odesolver);
      \end{scope}

      \begin{scope}[xshift=3.5cm]
        \node at (0,1) {Proposed method};
        \node (input) [io] {Initial state};
        \node (odesolver) [libcode, below of=input] {ODE Solver};
        \node (output) [io, below of=odesolver] {Final state};
        \draw [arrow] (input) -- (odesolver);
        \draw [arrow] (odesolver) -- (output);
        \node (rhs) [libcode, right of=odesolver, node distance=2.5cm]
          {General purpose\\ $M\,\text{grad}\,f$ routine};
        \draw [arrow] (odesolver) -- (rhs);
        \draw [arrow]  (rhs) -- (odesolver);
        \node (h) [usercode, right of=rhs, node distance=2.5cm] {$H$ function};
        \draw [arrow] (rhs) -- (h);
        \draw [arrow] (h) -- (rhs);
      \end{scope}
    \end{tikzpicture}
  \caption{Flowcharts for the standard and the proposed methods. Round cornered
    boxes represent library or general purpose code; right angled boxes contain
    code for the specific calculation at hand, which the user must
  supply.\label{fig:a}}
\end{figure}

%To make this discussion clearer, let us consider a specific problem, the
%classical harmonic oscilator~\cite{french:71}, and a specific ODE solver method,
%the explicit forward Euler method \cite{sir:2010}. The hamiltonian function for
%a classical oscilator with mass $m$ and restoring force constant $k$ is 
%\begin{equation}
%  H(q,p)=\frac{1}{2}kq^2+\frac{p^2}{2m}.
%\end{equation}
%Hamilton's equations~\eqref{eq:heqs} then read
%\begin{align}
%  \frac{dq}{dt}&=\frac{p}{m}&
%  \frac{dp}{dt}&=-kq,
%\end{align}
%wich are already in the form of eq.~\eqref{eq:numheqs}. Given initial values
%$q_0$ and $p_0,$ successive values can be estimated at times $t_1$, $t_2$, \ldots
%by iterating the formulas
%\begin{align}
%  q_{i+1}&=q_{i}+\frac{p_i}{m}\delta t_i\label{eq:efe1}\\
%  p_{i+1}&=p_{i}-kq_i\delta t_i\label{eq:efe2},
%\end{align}
%where $\delta t_i=t_{i+1}-t_i$. It is now easy to write a  program that
%implements this method, and this the usual procedure. With this paper, we
%consider a method where explicit expressions for the hamiltonian's partial
%derivatives in the equations of motion are replaced by numerical estimates. In
%our toy problem, that amounts to rewriting eqs~\eqref{eq:efe1}
%and~\eqref{eq:efe2} by
%\begin{align*}
%  q_{i+1}&=q_{i}+\numap{\frac{\partial H}{\partial p}}_i \delta t_i\\
%  p_{i+1}&=p_{i}-\numap{\frac{\partial H}{\partial q}}_i\delta t_i,
%\end{align*}
%where $\numap{X}_i$ means \emph{numerical approximation of $X$ at time $t_i$.}
%In order to make numerical aproximations to the partial derivatives of a
%function, one must evaluate it several times; at the very least, two evaluations
%are needed. For example, using the second order central difference method, the
%derivative is estimated as
%\begin{equation}\label{eq:cdiff}
%  \frac{\partial\psi}{\partial x} = 
%    \frac{\psi(x+\delta x)-\psi(x-\delta t)}{2\delta x}+
%  o(\delta x)^2,
%\end{equation}
%leading to approximation formulas that, for our example problem, read 
%\begin{align}
%  \numap{\frac{\partial H}{\partial p}}_i &=
%  \frac{H(q_i, p_i+\delta p)-H(q_i, p_i-\delta p)}{2\delta p}\\
%  \numap{\frac{\partial H}{\partial q}}_i &=
%  \frac{H(q_i+\delta q, p_i)-H(q_i-\delta q, p_i)}{2\delta q},
%\end{align}
%where $\delta q$ and $\delta p$ are suitably chosen discretization parameters.
%Other approximations have higher accuracy, at the cost of extra evaluations of
%the function to be derived.

%This need for repeated evaluation of the system's hamiltonian reduces the
%efficiency of this method. However, programming a procedure for the hamiltonian
%function is usually much simpler then doing it for the set of its partial
%derivatives.  Also, adjusting the syntax and semantics of the ODE solver and the
%hamiltonian procedure is likewise much simpler than when the partial derivatives
%are coded.  For small systems, using the hamiltonian is much simpler and faster.
%
%
%
\section{Simple implementation in Python and examples}
We made a simple implementation of this method in Python, using the standard
numerical extensions library NumPy~\cite{numpy:2011,numpy:2020}.  It consists of
a simple function that calls the hamiltonian function supplied by the user to
evaluate its partial derivatives, using the central difference formulas of
eqs.~\eqref{eq:cdiff}. This function is then supplied to NumPy's ODE solver
\texttt{solve\_ivp}.  For greater efficiency, it is wrapped in a class, whose
objects store initialization details on creation, like the hamiltonian's
parameters, the number of degrees of freedom, the values of the discretization
steps, etc. These objects are callable with signature adjusted to what the ODE
solver expects.  The whole class definition fits in a few dozen lines of code
and is very easy to use. It is available for download as free software at
github~\cite{nevh:2020}. The repository also includes a \texttt{jupyter
notebook} showcasing several example applications.

\section{Conclusion}

\section*{References}
\begin{thebibliography}{99}
  \bibitem{gol:1980} H. Goldstein, \textsl{Classical Mechanics.} Addison-Wesley
  (1980)
%
  \bibitem{fw:2003} 
    A.L.~Fetter, J.D.~Walecka,
    \textsl{Theoretical Mechanics of Particles and Continua.}
    Dover (2003)
%
  \bibitem{odepack}
  A.C. Hindmarsh, \textsl{ODEPACK, A Systematized Collection of ODE Solvers,} in
    Scientific Com\-puting, R. S. Stepleman et al. (eds.), North-Holland,
    Amsterdam, 1983 (vol. 1 of IMACS Transactions on Scientific Computation),
    pp. 55-64. 
%
  \bibitem{odeint}
  K. Ahnert and M. Mulansky, \textsl{Odeint - Solving Ordinary Differential
  Equations in C++,} AIP Conf. Proc. 1389, pp. 1586-1589 (2011);
doi:http://dx.doi.org/10.1063/1.3637934
%
  \bibitem{scipy}
  P. Virtanen et al, \textsl{SciPy 1.0: Fundamental Algorithms for Scientific
  Computing in Python}, Nature Methods 17, pp. 261--272 (2020)
%
  \bibitem{french:71} A.P.~French, \textsl{Vibration and Waves.} W.W. Norton \&
  Company (1971)
%
  \bibitem{sir:2010} S. \v{S}irca, M. Horvat, \textsl{Computational Methods in
  Physics.} Springer (2012)
%
  \bibitem{nr:2007} W.H.~Press, S.A.~Teukolsky, W.T.~Vetterling, B.P.~Flannery,
  \textsl{Numerical Recipes -- The Art of Scientific Computing,} 3rd Ed.
  Cambridge University Press (2007)
%
  \bibitem{numpy:2011} S.J.~van der Walt, S.C.~Colbert, G.~Varoquaux,
  \textsl{The NumPy array: a structure for efficient numerical computation,}
  Computing in Science \& Engineering, \textbf{13} no. 2, 22-30 (2011).
  doi: 10.1109/MCSE.2011.37
%
  \bibitem{numpy:2020} C.R.~Harris, K.J.~Millman, S.J.~van der Walt et. al.,
  \textsl{Array programming with NumPy,} Nature \textbf{585}, 357-362 (2020).
  https://doi.org/10.1038/s41586-020-2649-2
%
  \bibitem{nevh:2020} J.M.L.~Amoreira, L.J.M.~Amoreira,
  \texttt{https://github.com/ljmamoreira/nevh} (2020)

\end{thebibliography}
\end{document}